Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2021-04-21 manuscript/manuscript.Rmd
Suparna final edits

  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals' preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.

@@

In 2016, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory [@Roth:2016]. Governments across the globe have cracked down on civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017; @DPR:2015; @DPR:2016].[^NGOdef] Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations. Barriers to funding are especially pervasive and restrict the ability of NGOs to secure financial resources. Restrictive states may prevent the transfer of foreign funds to NGOs based on the origin and purposes of these funds.

As a result of these laws, public or official foreign aid channeled through international NGOs (INGOs) has decreased in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given the reduction of foreign aid following restricted legal environments abroad, INGOs may need to devote more resources to fundraise from individual donors.

Individual private donations are a critical source of INGO funding. While total private philanthropy[^phildef] in the United States has grown over time, it is notable that private giving to NGOs working on international affairs-focused issues are one of only two categories of organizations that have continued experiencing substantial growth in giving, reaching \$22.08 billion in donations in 2018 [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale (up to \$50) or micro-donations (between \$0.25–\$10) [@SaxtonWang:2014]. In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider their preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad?

A robust scholarship already examines the determinants of individual-level philanthropy to domestic causes. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor attributes and preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. Trust is an essential component for INGO sustainability—for many nonprofits, trust often translates into grants and donations [@coombs2007]. Geography often impedes donors' direct evaluation of their work and trust is one of the most influential criteria that private donors use when deciding which nonprofits to support [@prakash2010trust; @bekkers2003trust]. A lack of trust could be further fueled by legal trouble with the INGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in INGOs influence individuals' preferences to donate, especially when groups face crackdown?

We explore this using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate dozens of distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that an organization's target issue area and its relationship with its host government interact strongly with individuals' characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments. Instead, they prefer donating to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged organizations. Our findings imply that individual donors with high trust would remain supportive of INGOs when they face criticism and crackdown from foreign governments. On the other hand, government crackdown may further sour INGOs' least trusting donors.

@@

Second, knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Our paper shows that in readjusting their fundraising strategies towards individuals, framing, or the process through which actors present information to influence perceptions, is particularly important [@chong2007framing]. Framing affects their willingness to donate. However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help INGOs frame their appeals. In particular, our results show that INGOs may benefit from publicizing crackdown at least towards their most trusting donors.

Finally, while research in public policy examines giving to NGOs, the discipline of international relations tends to see philanthropy as less relevant [@youde2019]. However, philanthropy—particularly donations received by INGOs—can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, studying philanthropy changes our conceptions of individuals being passive recipients to active entities that are not merely objects of governance, and can thus shed light on the changing logic of global governance.

@@

Foreign aid to civil society—and NGOs in particular—has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. However, NGO-state relationships have also evolved—in the 1950s and 60s, international assistance channeled through NGOs aimed at producing socioeconomic change rather than political progress. But in the aftermath of the dissolution of the Soviet Union, many NGOs had the explicit goal of promoting democracy and bringing down authoritarian governments. Subsequently, many states see NGOs as a threat and seek to repress them.

More than 100 countries have proposed or enacted 244 measures restricting, repressing, or shutting down civil society since 2013 [@icnl:2021]. This has had a threefold impact on INGOs. First, it has led to a reduction in resources and grants available to INGOs. Second, restrictive NGO laws can affect INGO programming by preventing organizations from working in certain (contentious) issue areas. To deal with this, some groups have adapted by changing the nature of their programming to appear less threatening to the repressive countries. Finally, in the worst-case scenario, this has led to a loss in INGOs' access to target countries—often also associated with the use of violence against INGOs and their employees. We focus exclusively on the first dilemma—the reduction in INGO funding and possible responses to it.

@@

INGOs receive funds from a variety of public and private sources. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles in funding. Closing civic space has resulted in net losses of income to INGOs, particularly in official bilateral and multilateral aid [@DupuyPrakash:2018]

Legal crackdown has also affected how and where funds are allocated. Following the passage of anti-NGO laws, official donors move funds away from contentious causes such as human rights, media, corruption, advocacy, among others, towards tamer and more regime-compatible causes such as health, education [@ChaudhryHeiss:2018]. The latter trend has also been exacerbated by a preference among donors working in official aid and the democracy promotion establishment to gradually increase funding directed towards tame causes, at the expense of more contentious causes. This is because many tame causes can have easily quantifiable, measurable outputs, which makes it easier to show a program's effectiveness over time [@Bush:2015].

@@

We explore the interaction between individual and organizational characteristics using a conjoint experiment conducted with a nationally representative sample in the U.S. Conjoint analysis allows us to elicit donor preferences without respondents explicitly stating their exact preferences. This kind of experiment is commonly used in marketing to reveal consumer preferences for products and has become increasingly popular in political science and public policy [@KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy, where we can study the simultaneous effects of organizational characteristics and individual attributes on individuals' preferences to donate to an INGO.

@@

To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. Conjoint simulation offers useful analytic benefits. Since we are less concerned with the marginal effects of single donor characteristics and more interested in how constellations of individual and organizational attributes interact with each other, we can vary different combinations of donor and organization characteristics and hold other constant. This allows us to see the connections between donor trust and background and INGO organizational practices and find which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, and which donor characteristics are associated with the propensity to donate.

@@

An organization's relationship with its host government does influence market shares across levels of income, but the size of this difference depends on the magnitude of income. In general, personas prefer to donate to organizations with friendly relationships with their host governments and avoid those under crackdown. Wealthier donor personas are more likely to prefer donating to organizations that are criticized by their host governments. This falls in line with their preference for human rights organizations, which have more contentious missions and are more likely to be targeted. This support, however, evaporates as crackdown intensifies. High-income personas are the least likely demographic to prefer donating to NGOs under crackdown. Their lower income counterparts, on the other hand, are generally unfazed by NGOs' relationships with government, maintaining the same average market share across all three types of host country relationships. Patterns in preferences for issue area and funding sources remain constant across both methods of categorizing income—crackdown only has an effect on very high-income personas. Income thus appears to influence preferences particularly for NGOs facing crackdown abroad—wealthy donors are perhaps fair-weather friends that are more likely to pull their support when an organization's work becomes more difficult.

@@

Next, we next examine how differences in public affairs knowledge, experience, and political ideology influence a persona's propensity to donate to an organization. We vary personas along two dimensions: (1) political ideology, with semi-liberal and semi-conservative political leanings (2 and 5 on a 7-point scale), and (2) knowledge and experience with public affairs and the news, with personas that either follow national and international news and that have traveled abroad or personas that do not follow the news or travel.

@@

Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have between a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer instead to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

@@

Personas across both levels of social trust are most likely to donate to organizations with friendly relationships with their host governments. However, those with higher trust are less likely to be deterred by poor host-country relationships, while those with low trust tend to not donate to organizations facing criticism or crackdown abroad. A friendly relationship with a host government may be a manifestation of a belief that charities behave properly and exert a positive influence on society. Once an organization faces criticism or crackdown, however, support deteriorates rapidly among personas with low social trust, since such a crackdown could be perceived as a violation of the persona's trust in the charity—that is, the organization must be doing something wrong to deserve criticism. When the risk of a certain action is higher, people need higher levels of confidence or trust to engage in that action [@coleman1990]. When NGOs have goals that are harder to accomplish (as is the case under shrinking civic space), there is greater uncertainty that the donated money will meaningfully contribute to the intended cause.

Personas with high levels of trust, on the other hand, seem to stick with legally besieged organizations and show targeted groups the benefit of the doubt. The global crackdown on NGOs thus sours organizations' least trusting individual donors. Overall, our results support research that donors with high levels of trust are more active in giving than those with low levels of trust [@sargeant2008] and that trust especially matters when individuals donate to organizations working under great uncertainty and towards difficult goals [@Wiepking:2010].

@@

Our conjoint experiment raises two main concerns for inference: (1) prior knowledge and opinions about these organizations could influence respondents' choices, and (2) there are questions about the generalizability of these results outside the American context. We explore both below.

@@

We also failed to find any systematic evidence that the Oxfam scandal influenced perceptions of the nonprofit sector in the U.S., likely because the story had far less news coverage in the U.S. Many individual donors have a preference to donate locally and they lack sufficient knowledge to make an informed decision about donating internationally. @KnowlesSullivan:2017, conducting a field experiment in New Zealand, found that only a small number of participants in the study stated that not giving internationally was driven by mistrust. Rather, lack of trust towards overseas NGOs, particularly aid and developmental agencies often results due to donors perceiving INGOs as having low personal relevance [@FaulknerTruongRomaniuk:2015]. Overall, the literature suggests that for INGO fundraising, lack of awareness is a much larger issue than specific scandals or other knowledge of specific organizational characteristics that might reduce donor trust and confidence.

@@

The global crackdown on civil society has adversely impacted the operations of NGOs. It has not just resulted in net losses of income from official sources to INGOs, it has also affected how and where funds are allocated. We argue that individual donors can be useful in sustaining the operations of groups working in challenging environments. Individual donors do not use the same performance-based metrics as official aid donors and use heuristics to simplify their decision-making. The role of trust may be important in understanding donations preferences towards INGOs working in challenging contexts. While trust is a difficult construct to define, the nonprofit literature conceptualizes trust as a key driver of commitment, which is shown to impact both intention to donate as well as actual donation behavior. With repressive environments and NGOs working on more difficult goals, a lack of trust could be further fueled by legal trouble with the NGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in NGOs influence individuals' preferences to donate, especially when groups face crackdown?

Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find several important trends. Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to emergency response and refugee relief organizations with unfriendly relationships with their host governments. Personas with low social trust tend to be wary of negative host-country relationships and strongly prefer to donate to organizations that are friendly with their host governments, while personas with high social trust are more willing to stick with legally besieged organizations. This implies that donors who are the most trusting could remain supportive when nonprofits face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors. Our findings lend support to research showing that individuals who trust NGOs are more likely to continue their monetary support of NGOs in the face of difficulties and under great uncertainty [@dwyer1987; @Wiepking:2010].

>>>>

  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income for INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors. Rather, trust can be an important component of individual donor support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals' preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.

@@

In 2016, Human Rights Watch claimed that civil society was under more aggressive attack than at any time in recent memory [@Roth:2016]. Globally, governments are repressing civil society organizations (CSOs) and non-governmental organizations (NGOs)—a phenomenon known as "closing civic space" [@CarothersBrechenmacher:2014; @CIVICUS:2017; @DPR:2015; @DPR:2016].[^NGOdef] Legal restrictions, or what we refer to as legal crackdowns, are a core part of these efforts. These crackdowns create barriers to entry, funding, and advocacy for NGOs in an effort to control, obstruct, and repress these organizations. Barriers to funding are especially pervasive and restrict the ability of NGOs to secure financial resources. Restrictive states may prevent the transfer of foreign funds to NGOs based on the origin and purposes of these funds.

Due to these laws, official foreign aid channeled through international NGOs (INGOs) has decreased in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given the reduction of foreign aid following restricted legal environments abroad, INGOs may need to fundraise more from individual donors.

Individual private donations are an important source of INGO funding. While total private philanthropy[^phildef] in the United States has grown over time, it is notable that private giving to NGOs working on international affairs-focused issues are one of only two categories of organizations that have continued experiencing substantial growth in giving, reaching \$22.08 billion in donations in 2018 [@givingusa:2019].[^intlaffairs] Most of this growth is driven by individuals making small-scale (up to \$50) or micro-donations (between \$0.25–\$10) [@SaxtonWang:2014]. In 2017, 70% of donations to international affairs NGOs came from individuals, rising to 86% in 2018 [@givingusa:2019]. But relying on individual donors means that NGOs must consider their preferences when fundraising. Which individual donor characteristics and attributes are most salient for INGOs interested in maximizing philanthropy at a time when they are facing increasingly hostile environments abroad?

A robust scholarship examines the determinants of individual-level philanthropy to domestic causes. But the dynamics of *international* giving to NGOs are under-examined in the literature on individual donor preferences.[^exceptions] Motivating individuals to donate to international causes is hard as the number of recipients are large and far removed from the donor [@CasaleBaumann:2015, 100]. Trust is an essential component for INGO sustainability—for many nonprofits, trust often translates into grants and donations [@coombs2007]. Geography often impedes donors' direct evaluation of their work and trust is one of the most influential criteria that private donors use when deciding which nonprofits to support [@prakash2010trust; @bekkers2003trust]. A lack of trust could be further fueled by legal trouble with the INGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in INGOs influence individuals' preferences to donate, especially when groups face crackdown?

We explore this using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity. We adapt methods used in marketing to generate distinct simulated personas that represent typical donor profiles, and we vary these personas along different dimensions of demographic, political, and social attributes. We find that an organization's target issue area and its relationship with its host government interact strongly with individuals' characteristics. Personas with high levels of trust in political and charitable institutions and who are actively engaged in charity and volunteerism prefer donating to human rights organizations that have friendly relationships with their host governments. In contrast, personas with low levels of social trust eschew both human rights organizations *and* any type of NGO with more contentious relationships with host governments. Instead, they prefer donating to NGOs with friendly host-country relationships and working on less contentious issues like emergency response and refugee relief. Personas with higher trust are more comfortable donating to legally besieged organizations. Our findings imply that individual donors with high trust would remain supportive of INGOs when they face criticism and crackdown from foreign governments. On the other hand, government crackdown may further sour INGOs' least trusting donors.

@@

Second, knowledge of demographic traits and its impact on funding can help nonprofits craft more effective and targeted marketing strategies [@Rajan:2009, 414]. Our paper shows that in readjusting their fundraising strategies towards individuals, framing, or the process through which actors present information to influence perceptions, is particularly important [@chong2007framing]. Framing affects individuals' willingness to donate. However, existing research is unclear about which frames are effective when making appeals to individual donors considering international philanthropy. Since our study looks at multiple organizational and host country factors, our results can help INGOs frame their appeals. Our results show that INGOs may benefit from publicizing crackdown at least towards their most trusting donors.

Finally, while research in public policy examines giving to NGOs, the discipline of international relations tends to see philanthropy as less relevant [@youde2019]. However, philanthropy can help us understand changes in the dynamics of global governance. Private individuals exercise power and influence in unique ways that differ from other non-state actors. As @youde2019 [p. 44] points out, studying philanthropy changes our conceptions of individuals as being passive recipients to active entities that are not merely objects of governance, and can thus shed light on the changing logic of global governance.

@@

Foreign aid to NGOs has steadily increased over time. Many government and foundation donors perceive NGOs as nimbler and less bureaucratic than governments. In countries with weak institutions and poor governance, where direct aid transfers to governments raise issues of misuse and bureaucratic inefficiency, donors specifically seek out NGOs for projects [@Dietrich2013]. However, NGO-state relationships have also evolved—in the 1950s and 60s, international assistance channeled through NGOs aimed at producing socioeconomic change rather than political progress. But in the aftermath of the dissolution of the Soviet Union, many NGOs had the explicit goal of promoting democracy and bringing down authoritarian governments. Subsequently, many states saw NGOs as a threat and sought to repress them.

More than 100 countries have proposed or enacted 244 measures restricting, repressing, or shutting down civil society since 2013 [@icnl:2021]. This has had a threefold impact on INGOs. It has led to a reduction in resources and grants available to INGOs. Restrictive NGO laws have affected INGO programming by preventing organizations from working in certain (contentious) issue areas. To deal with this, some groups have adapted by changing the nature of their programming to appear less threatening to the repressive countries. In the worst-case scenario, it has led to a loss in INGOs' access to target countries. We focus exclusively on the first dilemma—the reduction in INGO funding and possible responses to it.

@@

INGOs receive funds from a variety of public and private sources. Public funds—or traditional donor aid—channel money from official aid agencies through a variety of bilateral and multilateral institutions. Private sources include foundations, private corporations, and individuals. While INGOs have not been passive in responding to the crackdown on civil society, they face increasing obstacles in acquiring funding. Closing civic space has resulted in net losses of income to INGOs, particularly in official bilateral and multilateral aid [@DupuyPrakash:2018]

Legal crackdown has also affected how and where funds are allocated. Following the passage of anti-NGO laws, official donors move funds away from contentious causes such as human rights, media, corruption, advocacy, among others, towards tamer and more regime-compatible causes such as health, education [@ChaudhryHeiss:2018]. The latter trend has also been exacerbated by a preference among donors working in the official aid and democracy promotion establishment to gradually increase funding directed towards tame causes, at the expense of more contentious causes. This is because many tame causes can have easily quantifiable, measurable outputs, which makes it easier to show a program's effectiveness over time [@Bush:2015].

@@

We explore the interaction between individual and organizational characteristics using a conjoint experiment conducted with a nationally representative sample in the U.S. Conjoint analysis allows us to elicit donor preferences without respondents explicitly stating their preferences. This kind of experiment is commonly used in marketing to reveal consumer preferences for products and has become increasingly popular in political science and public policy [@KnudsenJohannesson:2019]. Because we are interested in exploring the complexity behind the decision to donate, we apply conjoint methods to the study of nonprofits and philanthropy, where we can study the simultaneous effects of organizational characteristics and individual attributes on individuals' preferences to donate to an INGO.

@@

To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. Conjoint simulation offers useful analytic benefits. Since we are less concerned with the marginal effects of single donor characteristics and more interested in how constellations of individual and organizational attributes interact with each other, we can vary different combinations of donor and organization characteristics and hold other constant. This allows us not only to see the connections between donor trust and INGO organizational practices and find which kinds of donors are more or less likely to consider specific organizational attributes when deciding to donate, but also which donor characteristics are associated with the propensity to donate.

@@

An organization's relationship with its host government does influence market shares across levels of income, but the size of this difference depends on the magnitude of income. In general, personas prefer to donate to organizations with friendly relationships with their host governments and avoid those under crackdown. Wealthier donor personas are more likely to prefer donating to organizations that are criticized by their host governments. This falls in line with their preference for human rights organizations, which have more contentious missions and are more likely to be targeted. This support, however, evaporates as crackdown intensifies. High-income personas are the least likely demographic to prefer donating to NGOs under crackdown. Their lower income counterparts, on the other hand, are generally unfazed by NGOs' relationships with governments, maintaining the same average market share across all three types of host country relationships. Patterns in preferences for issue area and funding sources remain constant across both methods of categorizing income—crackdown only has an effect on very high-income personas. Income thus appears to influence preferences particularly for NGOs facing crackdown abroad—wealthy donors are perhaps fair-weather friends that are more likely to withdraw their support when an organization's work becomes more difficult.

@@

Next, we examine how differences in public affairs knowledge, experience, and political ideology influence a persona's propensity to donate to an organization. We vary personas along two dimensions: (1) political ideology, with semi-liberal and semi-conservative political leanings (2 and 5 on a 7-point scale), and (2) knowledge and experience with public affairs and the news, with personas that either follow national and international news and that have traveled abroad or personas that do not follow the news or travel.

@@

Social trust plays a substantial role in determining simulated market share, however. Those who trust social and charitable institutions are very likely to prefer donating to human rights organizations, regardless of their public affairs knowledge or political leanings. In all circumstances, those with high levels of trust have a 7–10% chance of donating to a human rights organization, on average. Personas with low levels of trust, on the other hand, strongly avoid human rights organizations and prefer to donate to refugee and emergency response NGOs. The difference is substantial—a liberal persona with high levels of public affairs knowledge and experience will give to a human rights organization 10% of the time, on average if they have high levels of trust, but less than 3% of the time if they are less trusting. Trust in charities and institutions is a key factor in deciding to donate to more contentious human rights organizations.

@@

Personas across both levels of social trust are most likely to donate to organizations with friendly relationships with their host governments. However, those with higher trust are less likely to be deterred by poor host-country relationships, while those with low trust tend to not donate to organizations facing criticism or crackdown abroad. A friendly relationship with a host government may be a manifestation of a belief that NGOs act appropriately and exert a positive influence on society. Once an organization faces criticism or crackdown, however, support deteriorates rapidly among personas with low social trust, since such a crackdown could be perceived as a violation of the persona's trust in the charity—that is, the organization must be doing something wrong to deserve criticism. When the risk of a certain action is higher, people need higher levels of confidence or trust to engage in that action [@coleman1990]. When NGOs have goals that are harder to accomplish (as is the case under shrinking civic space), there is greater uncertainty that the donated money will meaningfully contribute to the intended cause.

Personas with high levels of trust, on the other hand, continue supporting legally besieged organizations and show targeted groups the benefit of the doubt. Crackdown on NGOs thus sours organizations' least trusting individual donors. Overall, our results support research that donors with high levels of trust are more active in giving than those with low levels of trust [@sargeant2008] and that trust especially matters when individuals donate to organizations working under great uncertainty and towards difficult goals [@Wiepking:2010].

@@

Our conjoint experiment raises two main concerns for inference: (1) prior knowledge and opinions about these organizations could influence respondents' choices, and (2) the generalizability of these results outside the American context. We explore both below.

@@

We also failed to find any systematic evidence that the Oxfam scandal influenced perceptions of the nonprofit sector in the U.S., likely because the story had far less news coverage in the U.S. Many individual donors have a preference to donate locally and they lack sufficient knowledge to make an informed decision about donating internationally. @KnowlesSullivan:2017, conducting a field experiment in New Zealand, found that only a small number of participants in the study stated that not giving internationally was driven by mistrust. Rather, lack of trust towards overseas NGOs often occurs due to donors perceiving INGOs as having low personal relevance [@FaulknerTruongRomaniuk:2015]. Overall, the literature suggests that for INGO fundraising, lack of awareness is a much larger issue than specific scandals or specific organizational characteristics that might reduce donor trust and confidence.

@@

The global crackdown on civil society has adversely impacted the operations of NGOs. It has not just resulted in net losses of income from official sources to INGOs, it has also affected how and where funds are allocated. We show that individual donors can be useful in sustaining the operations of groups working in challenging environments. Individual donors do not use the same performance-based metrics as official donors and use heuristics to simplify their decision-making. The role of trust may be important in understanding donations preferences towards INGOs working in challenging contexts. While trust is a difficult construct to define, the nonprofit literature conceptualizes trust as a key driver of commitment, which can impact both intention to donate as well as actual donation behavior. With repressive environments, a lack of trust could be further fueled by legal trouble with the NGO host government—a dynamic that individuals may not wish to enter with their dollars. How does trust in NGOs influence individuals' preferences to donate, especially when groups face crackdown?

Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find several important trends. Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences. Personas with high levels of social and charitable trust prefer donating to human rights organizations that have friendly relationships with their host governments, while personas with low trust are drawn more to emergency response and refugee relief organizations. Personas with low social trust tend to be wary of negative host-country relationships and strongly prefer to donate to organizations that are friendly with their host governments, while personas with high social trust are more willing to stick with legally besieged organizations. This implies that donors who are the most trusting could remain supportive when nonprofits face government criticism and crackdown. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings lend support to research showing that individuals who trust NGOs are more likely to continue their monetary support of NGOs in the face of difficulties and under great uncertainty [@dwyer1987; @Wiepking:2010].
